Bayesian Sparsification of Deep $\cplx$-valued networks

### 2020-02-07 paper/paper.tex
carved out the c-2r and 0.5c-r comparison

% model architectures
% discussion of TwoLayerDenseModel (TLDM) and SimpleConvModel (SCM)
The $\real$/$\cplx$ networks, studied in this experiment, have been chosen for the purpose 
of illustrating the compression and understanding the effects of experiment parameters.
\textit{SimpleConvModel} (SCM) is $2d$ ReLU convolutional networks with average pooling
($2\times 2$) after non-linearities with two convolutions (kernel $5\times 5$, stride $1$)
with filters $n_1 \to n_2$ followed by dense layers $n_3 \to n_\mathtt{out}$. The second
network, \textit{TwoLayerDenseModel} (TLDM), is a wide dense network $n_1 \to n_\mathtt{out}$.
The values of $n_\cdot$ depend on the input features and are summarized in table~\ref{tbl:mnist-like-model-spec}.

\begin{table}[hb]
  \centering
  \caption{Architecture parameters for models studied on MNIST-like datasets.}
  \label{tbl:mnist-like-model-spec}
  \begin{tabular}{lrr}
    \toprule
    $(n_1, n_2, n_3)$  & $\real$-\textit{SCM} & $\cplx$-\textit{SCM} \\
    \midrule
    (\texttt{raw}) & $20,  50,  500$ & $10, 25, 250$ \\
    (\texttt{fft}) & $40, 100, 1000$ & $20, 50, 500$ \\
  %   \bottomrule
  % \end{tabular}
  % \begin{tabular}{lrr}
    \toprule
    $(n_1)$  & $\real$-\textit{TLDM} & $\cplx$-\textit{TLDM} \\
    \midrule
    (\texttt{raw}) & $4096$ & $2048$ \\
    (\texttt{fft}) & $8192$ & $4096$ \\
    \bottomrule
  \end{tabular}
\end{table}

@@

% first experiment: 40-75-40 with fast scheduler
First we compare the $\real$-valued networks against their $\cplx$-valued counterparts with
approximately half as many parameters as suggested by \citet{monning_evaluation_2018}. In
this experiment stages last for $40$, $75$ and $40$ epochs, respectively, and the base learning
rate is scaled after $5$, $10$, and $20$-th epoch by $\tfrac1{10}$, $\tfrac1{20}$ and $\tfrac1{100}$,
respectively.
%
Plots~\ref{fig:mnist-like__trade-off} depict performance-compression curve on MNIST-like
datasets for the studied models and methods from sec.~\ref{ssub:variational_dropout} (VD)
and~\ref{ssub:empirical_bayes} (ARD). Besides that we tried \texttt{raw} or \texttt{fft}
input features, varied $
  C \in \{\tfrac3{2^{k+1}}\colon k=1..16\}
$, and repeated the experiment $4$ times.

% observation: raw features -- alright

fft features with wide dense network exhibit improving performance with the compression rate
up to a certain limit.

% discuss the plots describing how the test-performance grows with compression and then drops if the model is too sparse
% which is better ARD or VD
It is worth pointing out that the prior used in ARD models offers slightly more favourable
compression-accuracy values, matching the established perception that log-uniform prior
is too harsh for sparsification tasks \citep{kharitonov_variational_2018}.

Nevertheless, it is clear form the results depicted in \ref{where}, that VD and ARD
methods yield very similar compression-performance frontiers.

% second experiment: use scheduler Trabelsi et al. (2018) and show them plotz!
In the second experiment we use a much slower learning rate scheduler and study only the 
VD method.

% hinting at overfitting/regularization effects of masking and sparsification

% hopeful conclusion
\todo{The numerical experiments on MNIST-like datasets} demonstrate, that models using fourier
features perform poorly in comparison to models using raw image input.

@@

% mention the same experimental setup as in MNIST
Experiments on CIFAR10 dataset are set up similarly to MNIST-like experiment, except
that we study VGG16 model \citep{simonyan_very_2015} and its $\cplx$-counterpart. Each
network is trained stagewise for $20$, $40$, and $20$ epochs on the full training split
of $50k$ colour $32\times32$ images. We use only raw colour features (\texttt{raw}), since
the results of the experiments on MNIST left us sceptical about the utility of the Fourier
features for image recognition. We use fast learning rate scheduler from sec.~\ref{sub:mnist_like_datasets}
and apply random crop and horizontal flips as augmentation.

@@

% draw some conclusions
The halved $\cplx$ network significantly underperforms in comparison with its $\real$ counterpart

@@

In this section we derive the exact gradient of \eqref{eq:improper-kl-div-real}. Even
though, ultimately, the divergence involves a hypergeometric function \todo{which one},
and its gradient requires evaluating the Gauss error function, the analysis provides
independent verification of the approximation, proposed by \citet{molchanov_variational_2017}.
The logic of the derivation follows \citet{lapidoth_capacity_2003}.
% \citet{pav_moments_2015} correctly notice that \cite[p. 2466]{lapidoth_capacity_2003} has a missing $\log{2}$ term.

@@

>>>>

% model architectures
% discussion of TwoLayerDenseModel (TLDM) and SimpleConvModel (SCM)
The $\real$/$\cplx$ networks, studied in this experiment, have been chosen for the purpose 
of illustrating the compression and understanding the effects of experiment parameters.
\textit{SimpleConvModel} (SCM) is $2d$ ReLU convolutional networks with average pooling
($2\times 2$) after non-linearities with two convolutions (kernel $5\times 5$, stride $1$)
with filters $n_1 \to n_2$ followed by dense layers $n_3 \to n_\mathtt{out}$. The second
network, \textit{TwoLayerDenseModel} (TLDM), is a wide dense network $n_1 \to n_\mathtt{out}$.
For SCM model $n_1, n_2, n_3$ are $20, 50, 500$, respectively, and for TLDM model $
  n_1 = 4096
$ regardless of their type ($\real$ or $\cplx$), although \citet{monning_evaluation_2018}
argues that this renders models incomparable.

@@

% first experiment: 40-75-40 with fast scheduler
In this experiment stages last for $40$, $75$ and $40$ epochs, respectively, and the base
learning rate of ${10}^{-3}$ is reduced after the $10$-th epoch to ${10}^{-4}$.
%
Plots~\ref{fig:mnist-like__trade-off__kmnist} and \ref{fig:mnist-like__trade-off__emnist}
depict samples from the performance-compression curve on KMNIST and EMNIST-Letters datasets
for the studied models and VD compression method (sec.~\ref{ssub:variational_dropout}).
Besides that we tried \texttt{raw} or \texttt{fft} input features, varied $
  C \in \{\tfrac3{2^{k+1}}\colon k=1..16\}
$, and repeated the experiment $5$ times.

We make a couple of observations based on these results. First, Fourier features appear to
perform on par with raw colour channel data in the convolutional networks across most
compression rates, but require high compression to make wide shallow dense network work
as well. The upward trend in manifests itself in MNIST and Fashion MNIST experiments as well
(see appendix~\ref{sub:other_mnist_like_datasets}). Finally, overall $\cplx$ variational
dropout methods seems to compress network adequately well, without much loss in performance.

@@

% mention the same experimental setup as in MNIST
Experiments on CIFAR10 dataset are set up similarly to MNIST-like experiment, except
that we study VGG16 model \citep{simonyan_very_2015} and its $\cplx$-counterpart and
replicate each experiment only twice. Each network is trained stagewise for $20$, $40$,
and $20$ epochs on the full training split of $50k$ colour $32\times32$ images. We use
only raw colour features (\texttt{raw}). We use fast learning rate scheduler and apply
random crop and horizontal flips as augmentation. The scheduler scales the base learning
rate after $5$, $10$, and $20$-th epoch by $\tfrac1{10}$, $\tfrac1{20}$ and $\tfrac1{100}$,
respectively.

@@

% draw some conclusions
Figure~\ref{fig:figure__cifar10__trade-off} shows that although the prior used in ARD method
offers slightly less compression, it makes up for it by equally slightly better accuracy.
Furthermore, the results imply that that the halved $\cplx$ network visibly underperforms
in comparison with its $\real$ counterpart, although in MNIST-like experiment $\cplx$-valued
network with the same inner feature dimensions performed on close to $\real$-valued, which
is half the size of the former.

@@

In this section we derive the exact gradient of \eqref{eq:improper-kl-div-real}. Even
though, ultimately, the divergence involves a hypergeometric function, and its gradient
requires evaluating the Gauss error function, the analysis provides independent verification
of the approximation, proposed by \citet{molchanov_variational_2017}. The logic of the
derivation follows \citet{lapidoth_capacity_2003}.
% \citet{pav_moments_2015} correctly notice that \cite[p. 2466]{lapidoth_capacity_2003} has a missing $\log{2}$ term.

@@

\subsection{Other MNIST-like datasets} % (fold)
\label{sub:other_mnist_like_datasets}

We provide the figures of the MNIST-like experiments, omitted from the main paper
due to size limitations.

\begin{figure}[!h]
  \centering
  \begin{subfigure}[b]{1.\textwidth}  % imcl2019-style swears at this
    \centering
    \includegraphics[width=\linewidth]{../assets/figure__mnist-like__trade-off/legacy__VD__fashionmnist__fft__-0.5.pdf}
  \end{subfigure} \\%
  % \hfill
  \begin{subfigure}[b]{1.\textwidth}  % imcl2019-style swears at this
    \centering
    \includegraphics[width=\linewidth]{../assets/figure__mnist-like__trade-off/legacy__VD__fashionmnist__raw__-0.5.pdf}
  \end{subfigure}
  \caption{%
    The compression-accuracy trade-off on the letters part of the Fashion MNIST dataset.
    (\textit{top}) \texttt{fft} features, (\textit{bottom}) \texttt{raw} features.
  }
  \label{fig:mnist-like__trade-off__fashionmnist}
\end{figure}

\begin{figure}[!h]
  \centering
  \begin{subfigure}[b]{1.\textwidth}  % imcl2019-style swears at this
    \centering
    \includegraphics[width=\linewidth]{../assets/figure__mnist-like__trade-off/legacy__VD__mnist__fft__-0.5.pdf}
  \end{subfigure} \\%
  % \hfill
  \begin{subfigure}[b]{1.\textwidth}  % imcl2019-style swears at this
    \centering
    \includegraphics[width=\linewidth]{../assets/figure__mnist-like__trade-off/legacy__VD__mnist__raw__-0.5.pdf}
  \end{subfigure}
  \caption{%
    The compression-accuracy trade-off on the letters part of the MNIST dataset.
    (\textit{top}) \texttt{fft} features, (\textit{bottom}) \texttt{raw} features.
  }
  \label{fig:mnist-like__trade-off__mnist}
\end{figure}

% subsection other_mnist_like_datasets (end)
